### 2019-12-27 overview.tex
cleaned up some mess from yesterday

In this chapter, we give a brief theoretical background of the topics studied.

mention the abuse of terms between droplets and particles.

The incompressible Navier-Stokes equation
\begin{subequations}
 \begin{equation}
   \nabla \cdot {\bm u} = 0,
  \label{eq:div-free}
 \end{equation}
 \begin{equation}
   \rho \bigg(\frac{\partial {\bm u}}{\partial t} + {\bm u} \cdot \nabla {\bm u} \bigg) = -\nabla p + \mu \nabla ^2  {\bm u} + {\bm f},
  \label{eq:NS}
 \end{equation}
\end{subequations}
governs the dynamics of Newtonian fluids (which is what we consider throughout the thesis).
where...

boundary conditions... between fluid and solid: continuity of the tangential velocity, \ie the no-slip condition (molecular slip ignored).
with an interface... surface tension... free surface... a boundary condition for the stress (see \cite{Batchelor} Sec.\ 3.3).

the Reynolds, Weber, and Froude numbers, defined as
\begin{equation}
  \begin{aligned}
    Re = \frac{\tilde{\rho} \tilde{U} \tilde{L}}{\tilde{\mu}},\quad \quad We = \frac{\tilde{\rho} \tilde{U}^2 \tilde{L}}{\tilde{\sigma}},\quad \quad Fr=\frac{\tilde{U}^2}{\tilde{g}\tilde{L}},      
  \label{eq:non-di-num}    
  \end{aligned}
\end{equation}
\noindent where $\tilde{U}$, $\tilde{L}$, $\tilde{\rho_1}$, $\tilde{\mu_1}$, and $\tilde{g}$ denote the reference dimensional velocity, length, density, dynamic viscosity, and gravitational acceleration. Note that $\rho_1=1$ and $\mu_1=1$ (i.e.\ we define fluid 1 as the reference fluid).
Perhaps also mention the Stouhal number.

@@

When small Reynolds number, Eq.\ \eqref{eq:NS} reduces to the Stokes equation,
\begin{equation}
   -\nabla p + \mu \nabla ^2  {\bm u} + {\bm f} = {\bm 0},
 \label{eq:Stokes}
\end{equation}
or, equivalently
\begin{equation}
   \nabla \cdot  {\bm \sigma} + {\bm f} = {\bm 0},
 \label{eq:Stokes1}
\end{equation}
where...

Stokes flow has the following symmetries and properties.
\begin{enumerate}
 \item \emph{Linearity}, which renders the applicability of superposition principle. For example, if $({\bm u}_1,p_1)$ is a solution to ${\bm f}={\bm f}_1$ and $({\bm u}_2,p_2)$ is a solution to ${\bm f}={\bm f}_2$, then $(\alpha{\bm u}_1+\beta {\bm u}_2,\alpha p_1 + \beta p_2)$ is a solution to ${\bm f}=\alpha{\bm f}_1 + \beta {\bm f}_2$.
 \item \emph{Reversibility}, which follows from the linearity and entails that, if the forcing changes from ${\bm f}$ to $-{\bm f}$, then the flow should reverse, \ie $({\bm u},p) \to (-{\bm u},-p)$. Note that, the simple reversibility symmetry can be very useful in deducing qualitative results without doing any calulation, see \eg the much celebrated note of \cite{Purcell1977}
 \item \emph{Stress equilibrium}, which states that any force exerted within the fluid is transmitted instantaneously to the boundary or, if there is no boundary, to infinity \citep{graham_2018}.
 \item \emph{Lorentz reciprocal relation}, which relates two solutions, $({\bm u}', {\bm \sigma}')$ and $({\bm u}'', {\bm \sigma}'')$ (they may differ by boundary conditions), of the Stokes equation by,
 \begin{equation}
  \small
  \begin{aligned}
   \int_V \bigg({\bm u}' \cdot (\nabla  \cdot {\bm \sigma}'') - {\bm u}'' \cdot (\nabla  \cdot {\bm \sigma}')\bigg) dV  & =\\
   \int_S \bigg({\bm u}' \cdot ({\bm n} \cdot {\bm \sigma}'') - {\bm u}'' \cdot ({\bm n} \cdot {\bm \sigma}')\bigg) dS, & 
  \end{aligned} \label{eq:lorentz-recip}
 \end{equation}
 where $V$ is some enclosed volume with surface $S$.
 \item \emph{Minimum dissipation principle}, which is a variational principle that asserts that the flow that minimizes energy dissipation rate subject to incompressibility and certain boundary conditions is the Stokes flow in that geometry.
\end{enumerate}

>>>>

This chapter summarizes the mathematical formulation and some basic theoretical ideas of microhydrodynamics and particle interactions.
By \emph{particle}, we loosely mean a solid or liquid body of micron to millimetre in size, which may also be referred to as \emph{droplet} for abuse of language.
As preliminaries, most of the contents are given without elaboration or proof. The interested reader is encouraged to read \cite{Batchelor, hb, ps, graham_2018} for greater details.

\bigskip

To begin with, recall the incompressible Navier-Stokes equation
\begin{subequations} \label{eq:Navier-Sotkes}
 \begin{equation}
   \nabla \cdot {\bm u} = 0,
  \label{eq:div-free}
 \end{equation}
 \begin{equation}
   \rho \bigg(\frac{\partial {\bm u}}{\partial t} + {\bm u} \cdot \nabla {\bm u} \bigg) = -\nabla p + \mu \nabla ^2  {\bm u} + {\bm f},
  \label{eq:NS}
 \end{equation}
\end{subequations}
relating the flow velocity $\bm u$ with the pressure $p$ in the presence of some force density $\bm f$, where $\rho$ and $\mu$ are the density and dynamic viscosity of the fluid, respectively.
Eq.\ \eqref{eq:Navier-Sotkes}, together with appropiate boundary conditions, governs the dynamics of a Newtonian fluid (which is what we consider throughout the thesis) that satisfies
\begin{equation}
 \begin{aligned}
   {\bm \sigma} = -p {\bm I}+ \mu \bigg( \nabla {\bm u} + (\nabla {\bm u})^T \bigg),
 \end{aligned}
\end{equation}
where $\bm \sigma$ is the stress tensor, and $\bm I$ the identity matrix.

Depending on the bounding material, conditions on the velocity or stress may be prescribed as the boundary condition. In the case of a solid boundary, it is usually the \emph{no-slip} velocity condition; while in the case of a liquid boundary, the following stress condition shall be satisfied,
\begin{equation} \label{eq:stress-bc}
  {\bm n} \cdot ({\bm \sigma}_+ - {\bm \sigma}_- )  = \gamma \kappa {\bm n} - \nabla \gamma,
\end{equation}
where $\bm n$ denotes the unit normal vector across the liquid interface (from $-$ to $+$), $\gamma$ the surface tension coefficent, and $\kappa$ the curvature of the interface (\eg $\kappa=2/a$ for a sphere of radius $a$).

Eq.\ \eqref{eq:NS} can be further non-dimensionalized, using the reference length $L$, velocity $U$, time $T$, pressure $P \equiv \mu U/L$, and force density $F \equiv P/L$ scales, into
\begin{equation}
   \textrm{Re} \bigg(\textrm{Sr}^{-1} \frac{\partial \hat{\bm u}}{\partial \hat{t}} + \hat{\bm u} \cdot \hat{\nabla} \hat{\bm u} \bigg) =
   -\hat{\nabla} \hat{p} + \hat{\nabla} ^2  \hat{\bm u} + \hat{\bm f},
  \label{eq:NS-nondi}
\end{equation}
where the hat signifies dimensionless quantities, and Re and Sr represent the Reynolds and Strouhal numbers, defined as
\begin{equation}
  \begin{aligned}
    \textrm{Re} = \frac{\rho UL}{\mu},\quad \quad \textrm{Sr} = \frac{UT}{L}.
  \end{aligned}
\end{equation}
A capilary number can also be defined to non-dimensionalize the surface tension,
\begin{equation}
  \begin{aligned}
    \textrm{Ca} = \frac{\mu U}{\gamma}.
  \end{aligned}
\end{equation}

@@

When both Re and ReSr$^{-1}$ are small, Eq.\ \eqref{eq:NS-nondi} reduces to the Stokes equation,
\begin{equation}
   -\nabla p + \mu \nabla ^2  {\bm u} + {\bm f} = {\bm 0},
 \label{eq:Stokes}
\end{equation}
or, equivalently
\begin{equation}
   \nabla \cdot  {\bm \sigma} + {\bm f} = {\bm 0},
 \label{eq:Stokes1}
\end{equation}
which are written in the dimensional form.

Physically, the Stokes equation describes systems that are characterized by small length scales or are in slow motions typically encountered in microfluidic or biological applications.
For instance, the Reynolds number based on a 10 $\mu$m droplet traveling at a speed of 10 $\mu$m/s in water ($\rho=10^3$ kg/m$^{3}$, $\mu=10^{-3}$ Pa$\cdot$s) is $10^{-4}$;
and if there is no imposed frequency, the Strouhal number will be 1, thus both Re $\ll 1$ and ReSr$^{-1}$ $\ll 1$ are safely satisfied.

Mathematically, the removal of the nonlinear term in Eq.\ \eqref{eq:NS} can dramatically simplify the solution.
In general, the Stokes flow has the following symmetries and properties.

\begin{enumerate}
 \item \emph{Linearity}, which makes solutions to the Stokes equation superimposible. For example, if $({\bm u}_1,p_1)$ is a solution to ${\bm f}={\bm f}_1$ and $({\bm u}_2,p_2)$ is a solution to ${\bm f}={\bm f}_2$, then $(\alpha{\bm u}_1+\beta {\bm u}_2,\alpha p_1 + \beta p_2)$ is a solution to ${\bm f}=\alpha{\bm f}_1 + \beta {\bm f}_2$.
 \item \emph{Reversibility}, which follows from the linearity and entails that, if the forcing changes from ${\bm f}$ to $-{\bm f}$, then the flow should reverse, \ie $({\bm u},p) \to (-{\bm u},-p)$. While seemingly redundant, the reversibility symmetry can be very powerful in making \emph{reductio ad absurdum} type-of arguments, see \eg the classical note of \cite{Purcell1977}
 \item \emph{Stress equilibrium}, which states that any force exerted within the fluid is transmitted instantaneously to the boundary or, if there is no boundary, to infinity \citep{graham_2018}.
 \item \emph{Lorentz reciprocal identity}, which relates two solutions, $({\bm u}', {\bm \sigma}')$ and $({\bm u}'', {\bm \sigma}'')$ (they may differ by boundary conditions), of the Stokes equation by,
 \begin{equation}
  \small
  \begin{aligned}
   \int_V \bigg({\bm u}' \cdot (\nabla  \cdot {\bm \sigma}'') - {\bm u}'' \cdot (\nabla  \cdot {\bm \sigma}')\bigg) dV  & =\\
   \int_S \bigg({\bm u}' \cdot ({\bm n} \cdot {\bm \sigma}'') - {\bm u}'' \cdot ({\bm n} \cdot {\bm \sigma}')\bigg) dS, & 
  \end{aligned} \label{eq:lorentz-recip}
 \end{equation}
 where $V$ is some enclosed volume with surface $S$.
 \item \emph{Minimum dissipation principle}, which is a variational principle that asserts that the flow that minimizes the energy dissipation rate subject to incompressibility under certain boundary conditions is the Stokes flow under the same boundary conditions.
\end{enumerate}
